Contracts for unions and intersections are hard for Nickel

### 2021-05-31 paper.tex
Fix compilation + small typos

Intersections are less prevalent. An intersection is a value that belongs to
two types at the same time. A simple example is a number literal \nickel{1}
which could be accepted as both an \nickel{Int} and a \nickel{Float} without
having to perform conversions (Figure~\ref{fig:intersection-subytping-ex}). In
this example, the intersection \nickel{Int & Float} can be used as a subtype of
both \nickel{Int} and \nickel{Float}.

@@

Not all hope is lost, we could pass some extra shared state
to each one of the sub contracts and let them only fail if the other
side of the contract has failed as well, as shown in Figure
\ref{fig:inter-contracts-2}, where the shared state is
represented by a label \nickel{l}.
This idea is still quite reasonable, for instance, when executing
\nickel{g 1}, from Figure \ref{fig:inter-contracts}, the first contract
that gets checked is the argument of \nickel{Str -> Str} over the function parameter, so
something like \nickel{1 | Str}.
Of course, this fails, but before raising a violation, this contract could
check if the other side of the intersection (\nickel{Number -> Number}) has
also tried to raise negative blame on the same parameter ()\nickel{1}), since
it hasn't, the contract would return normally, and would mark itself as failed.

@@

On the other hand, if you were to do something like
\nickel{g true}, that would fail also negatively the contract
\nickel{Str -> Str} and, when checking if \nickel{true}
is a correct parameter for \nickel{Number -> Number} it would check
if the other side of the intersection has failed, and since it
has, a violation would be raised and the whole program would
correctly fail.

@@

It is worth noting, that intersection types do have this property,
since subtyping would allow to correctly type check a term like
\nickel{g 1} as both \nickel{Number -> Number} as well as
\nickel{Bool -> Num}, and then deducing that \nickel{g 1}
can be given the type \nickel{Num -> Num @& Bool -> Num}.

@@

The second set of rules is deterministic, and it is intended
to be simpler to implement.
However, this second set needs to perform operations
on the evaluation context, to decide whether or not to evaluate
a given contract.
\footnote{You can check more on this topic on Section 5.1 of
\cite{KeilThiemannUnionIntersection}}
For instance, in the program presented on Figure \ref{fig:inter-flat-contracts},
when using the value \nickel{(fun x => x) | Str -> Str)}
inside the execution of contract \nickel{C}, and applying it to \nickel{0},
the contract runtime would check the context of execution and it would
deduce that contract \nickel{Str -> Str does not apply in this case,
skipping it.

@@

Next, we can combine these two contexts into one, that essentially
has one single missing value, that will be used independently in both contexts
from Figure \ref{fig:valid-contexts}, as shown on Figure \ref{fig:invalid-context}.
This context is not a valid \nickel{Num -> Num} context, since it will apply
the value put on $\square$ (through \nickel{f}) to a non \nickel{Num} (\nickel{true}).
In a similar way, it is not a valid \nickel{Bool -> Bool} context either.
Therefore, according to the semantics from the Williams, Morris, and Wadler paper, as
shown on Figure \ref{fig:wmw-semantics}, it is not a valid
\nickel{(Num -> Num) & (Bool -> Bool)} context.

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let f = (*@$\square$@*) in
(f 3, f true) 
\end{lstlisting}
\caption{Invalid \nickel{(Num -> Num) & (Bool -> Bool)} context,
combining contexts from Figure \ref{fig:valid-contexts}}
\label{fig:invalid-context}
\end{figure}

>>>>

Intersections are less prevalent. An intersection is a value that belongs to
two types at the same time. A simple example is a number literal \nickel{1}
which could be accepted as both an \nickel{Int} and a \nickel{Float} without
having to perform conversions (Figure~\ref{fig:intersection-subytping-ex}). In
this example, the intersection \nickel{Int @& Float} can be used as a subtype of
both \nickel{Int} and \nickel{Float}.

@@

Not all hope is lost, we could pass some extra shared state
to each one of the sub contracts and let them only fail if the other
side of the contract has failed as well, as shown in Figure
\ref{fig:inter-contracts-2}, where the shared state is
represented by a label \nickel{l}.
This idea is still quite reasonable, for instance, when executing
\nickel{g 1}, from Figure \ref{fig:inter-contracts}, the first contract
that gets checked is the argument of \nickel{Str -> Str} over the function parameter, so
something like \nickel{1 | Str}.
Of course, this fails, but before raising a violation, this contract could
check if the other side of the intersection (\nickel{Num -> Num}) has
also tried to raise negative blame on the same parameter ()\nickel{1}), since
it hasn't, the contract would return normally, and would mark itself as failed.

@@

On the other hand, if you were to do something like
\nickel{g true}, that would fail also negatively the contract
\nickel{Str -> Str} and, when checking if \nickel{true}
is a correct parameter for \nickel{Num -> Num} it would check
if the other side of the intersection has failed, and since it
has, a violation would be raised and the whole program would
correctly fail.

@@

It is worth noting, that intersection types do have this property,
since subtyping would allow to correctly type check a term like
\nickel{g 1} as both \nickel{Num -> Num} as well as
\nickel{Bool -> Num}, and then deducing that \nickel{g 1}
can be given the type \nickel{Num -> Num @& Bool -> Num}.

@@

The second set of rules is deterministic, and it is intended
to be simpler to implement.
However, this second set needs to perform operations
on the evaluation context, to decide whether or not to evaluate
a given contract.
\footnote{You can check more on this topic on Section 5.1 of
\cite{KeilThiemannUnionIntersection}}
For instance, in the program presented on Figure \ref{fig:inter-flat-contracts},
when using the value \nickel{(fun x => x) | Str -> Str)}
inside the execution of contract \nickel{C}, and applying it to \nickel{0},
the contract runtime would check the context of execution and it would
deduce that contract \nickel{Str -> Str} does not apply in this case,
skipping it.

@@

Next, we can combine these two contexts into one, that essentially
has one single missing value, that will be used independently in both contexts
from Figure \ref{fig:valid-contexts}, as shown on Figure \ref{fig:invalid-context}.
This context is not a valid \nickel{Num -> Num} context, since it will apply
the value put on $\square$ (through \nickel{f}) to a non \nickel{Num} (\nickel{true}).
In a similar way, it is not a valid \nickel{Bool -> Bool} context either.
Therefore, according to the semantics from the Williams, Morris, and Wadler paper, as
shown on Figure \ref{fig:wmw-semantics}, it is not a valid
\nickel{Num -> Num @& Bool -> Bool} context.

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let f = (*@$\square$@*) in
(f 3, f true) 
\end{lstlisting}
\caption{Invalid \nickel{Num -> Num @& Bool -> Bool} context,
combining contexts from Figure \ref{fig:valid-contexts}}
\label{fig:invalid-context}
\end{figure}
